### 2021-03-14 doc/manuscript/Dow_manuscript_Nature.Rmd
tweaks to results and discussion

**Growth milestones were achieved earlier at SCBI than at Harvard Forest (average difference: 6-10 days) and differed between ring- and diffuse-porous species (Fig. `r growth_curves_all`, Table `r mean_growth_parameters_table`).**
Across all trees and years, $DOY_{25}$ occurred on day 139 (May 20) at SCBI and day 148 (May 29) at Harvard Forest, while $DOY_{75}$ occurred on day 185 (July 5) at SCBI and day 193 (July 13) at Harvard Forest.
Consistent with the results of Ref[D'Orangeville], ring-porous species began growing earlier, reaching the $DOY_{25}$ benchmark earlier (by `r 154-123` days at SCBI and `r 164-132` at Harvard Forest), and their growth was spread out over a longer growing season (average $L_{PGS}$ 21 and 19 days longer at SCBI and Harvard Forest, respectively). 
Growing season length was similar across sites, with $L_{PGS}$ being, on average, only two days longer at SCBI for ring-porous species and <1 day longer for diffuse-porous species.

@@

The correlation of $DOY_{25}$ to $T_{min}$ was less consistent than that of $T_{max}$ (Figs. `r all_climwin_SI_figs`). 
At SCBI, patterns were quite similar to those described above for $T_{max}$, with similar timing and strengths of correlations across the two temperature variables (Figs. `r climwin_rp_DOY25_SCBI` - `r climwin_dp_DOY25_SCBI`).
In contrast, at Harvard Forest, where we had only six years of data, *climwin* selected windows, both > 5 weeks prior to $DOY_{25}$, had positive correlations between $DOY_{25}$ and $T_{min}$.
However, for both ring- and diffuse- porous species, average $T_{min}$ of the selected period was below freezing (<0°C) in all years, suggesting that these are sporadic correlations that are unlikely to be biologically meaningful.

@@

For both ring- and diffuse- porous species at both sites, warmer spring $T_{max}$ resulted in earlier achievement of growth milestones (Fig. `r schematic_summary`, Figs. `r pheno_Tsensitivity`).
Specifically, $DOY_{25}$, $DOY_{50}$, $DOY_{75}$, and $DOY_{g.max}$ at both sites all decreased significantly with increasing $T_{max}$ during the critical temperature window (Figs. `r schematic_summary`, `r pheno_Tsensitivity`, `r DOYtiming`).

@@

@@

We found that warmer spring temperatures in temperate deciduous forests cause earlier tree growth but have little effect on annual woody productivity. Using dendroband data, we found that the peak growing season is generally advancing as the critical spring windows are warming but having little, or negative effect, on $L_{pgs}$. $g_{max}$ and $\Delta DBH$ were virtually unchanged when this period warmed, but $DOY_{g_{max}}$ shifted back almost in unison with $DOY_{50}$. The lack of influence Spring TMAX exudes on $\Delta DBH$ found with the dendroband data agrees with patterns observed in tree cores. Annual tree growth depends more on peak growing season conditions than early season advancement.

Although the number of years analyzed here is limited (10 years at SCBI, 6 years at HF), the signal is clear and consistent: for both species groups at both sites, warmer temperatures during virtually any part of the spring led to significantly earlier achievement of growth milestones (Figs. `r schematic_summary`-`r `r DOYtiming``).
This indicates that the effect of $T_{max}$ on the timing of growth ($DOY_{25}$, $DOY_{50}$, $DOY_{75}$,and $DOY_{g_{max}}$) is not subtle; 
indeed, growth milestones advanced by 1-8 days per degree Celsius during the critical temperature window.
This is consistent with previous studies showing that ... *(find/cite some statistics on how rapidly leaf phenology or start of C sequestration advance with T)*

@@

Our findings contribute to a broader picture of the influence of spring temperatures on forest ecosystem function. 
Warm spring temperatures result in earlier leaf-out (fig., REFS), an earlier start to net C sequestration (REFS), and--as shown here--earlier initiation of stem growth (Figs. `r schematic_summary`- `r DOYtiming`).
Yet, an earlier start to the growing season does not imply a longer growing season (Figs. `r schematic_summary`, `r pheno_Tsensitivity`) [@zani_increased_2020].
Rather, growth finishes sooner-- presumably because of *C sink limitation* [@zani_increased_2020].

As climate change accelerates and spring temperatures become increasingly warmer, growing seasons will start earlier, but this is unlikely to provide the increased carbon sequestration that is anticipated in most models (Figs. `r schematic_summary`, `r pheno_Tsensitivity`, `r tree_rings`)[@zani_increased_2020].
The implication is that temperate deciduous forests are likely to sequester less CO~2~ than most models suggest, implying that models are underestimating future atmospheric CO~2~ overall.

>>>>

**Growth milestones were achieved earlier at SCBI than at Harvard Forest (average difference: 6-10 days) and differed between ring- and diffuse-porous species (Fig. `r growth_curves_all`, Table `r mean_growth_parameters_table`).**
Across all trees and years, the mean $DOY_{25}$ occurred on day 139 (May 20) at SCBI and day 148 (May 29) at Harvard Forest, while $DOY_{75}$ occurred on day 185 (July 5) at SCBI and day 193 (July 13) at Harvard Forest.
Consistent with the results of Ref[D'Orangeville], ring-porous species began growing earlier, reaching the $DOY_{25}$ benchmark earlier (by `r 154-123` days at SCBI and `r 164-132` at Harvard Forest), and their growth was spread out over a longer growing season (average $L_{PGS}$ 21 and 19 days longer at SCBI and Harvard Forest, respectively). 
Growing season length was similar across sites, with $L_{PGS}$ being, on average, only two days longer at SCBI for ring-porous species and less than one day longer for diffuse-porous species.

@@

The correlation of $DOY_{25}$ to $T_{min}$ was less consistent than that of $T_{max}$ (Figs. `r all_climwin_SI_figs`). 
At SCBI, patterns were quite similar to those described above for $T_{max}$, with similar timing and strengths of correlations across the two temperature variables (Figs. `r climwin_rp_DOY25_SCBI` - `r climwin_dp_DOY25_SCBI`).
In contrast, at Harvard Forest, where we had only six years of data, *climwin* selected windows greater than five weeks prior to $DOY_{25}$, had positive correlations between $DOY_{25}$ and $T_{min}$.
However, for both ring- and diffuse- porous species, average $T_{min}$ of the selected period was below freezing (<0°C) in all years, suggesting that these are sporadic correlations that are unlikely to be biologically meaningful.

@@

For both ring- and diffuse- porous species at both sites, warmer spring $T_{max}$ resulted in earlier achievement of growth milestones (Fig. `r schematic_summary`, Figs. `r pheno_Tsensitivity`).
Specifically, $DOY_{25}$, $DOY_{50}$, $DOY_{75}$, $DOY_{g.max}$, leaf greenup, and mid-greenup, at both sites all decreased significantly with increasing $T_{max}$ during the critical temperature window (Figs. `r schematic_summary`, `r pheno_Tsensitivity`, `r DOYtiming`).

@@

Ring- and diffuse- porous species were pooled together at each site for leaf phenology measures. Day of year of greenup occurred 21 days earlier at SCBI and 7 days earlier at Harvard Forest when comparing the year with the warmest and coldest critical windows in the analysis (Fig. 3). Mid-greenup timing also shifted forward at both sites when spring temperatures were higher but less so than greenup. Senescence occurred earlier at Harvard forest by 12 days in the warm year while SCBI saw a 5 day delay of senescence during the warm year.

@@

We found that warmer spring temperatures in temperate deciduous forests cause earlier tree growth but have little effect on annual woody productivity. Using dendroband data, we found that the peak growing season is generally advancing as the critical spring windows are warming. Despite this shift, growth variables $L_{pgs}$. $g_{max}$ and $\Delta DBH$ were virtually unchanged when this period warmed, but $DOY_{g_{max}}$ shifted back almost in unison with $DOY_{50}$. The lack of influence Spring TMAX exudes on $\Delta DBH$ found with the dendroband data agrees with patterns observed in tree cores, which show that annual tree growth depends more on peak growing season conditions than early season advancement.

Although the number of years analyzed here is limited (10 years at SCBI, 6 years at HF), the signal is clear and consistent: for both species groups at both sites, warmer temperatures during virtually any part of the spring led to significantly earlier achievement of growth milestones (Figs. `r schematic_summary`-`r `r DOYtiming``).
This indicates that the effect of $T_{max}$ on the timing of growth ($DOY_{25}$, $DOY_{50}$, $DOY_{75}$,and $DOY_{g_{max}}$) is not subtle. 
Indeed, growth milestones advanced by 1-8 days per degree Celsius during the critical temperature window.
This is consistent with previous studies showing that ... *(find/cite some statistics on how rapidly leaf phenology or start of C sequestration advance with T)*

@@

Our findings contribute to a broader picture of the influence of spring temperatures on forest ecosystem function. 
Warm spring temperatures result in earlier leaf-out (fig., REFS), an earlier start to net C sequestration (REFS), and, as shown here, earlier initiation of stem growth (Figs. `r schematic_summary`- `r DOYtiming`).
Yet, an earlier start to the growing season is not associated with a longer growing season (Figs. `r schematic_summary`, `r pheno_Tsensitivity`) [@zani_increased_2020]
because growth finishes sooner under these conditions, presumably because of *C sink limitation* [@zani_increased_2020].

As climate change accelerates and spring temperatures become increasingly warmer, growing seasons will start earlier, but this is unlikely to provide the increased carbon sequestration that is anticipated in most models (Figs. `r schematic_summary`, `r pheno_Tsensitivity`, `r tree_rings`)[@zani_increased_2020].
Because of this, temperate deciduous forests are likely to sequester less CO~2~ than most models suggest, implying that these models are underestimating future overall atmospheric CO~2~.
